Carrier phase and amplitude estimation for phase shift keying with unequal error protection

### 2012-12-03 other/uep/paper.tex
added to uep.  Needs code, should be fine now.

The sequences in $S$ can be enumerated as follows.  Let 
\[
T = \bigcup_{m\in G}\bigcup_{i \in D_m} \{ (z_i + \frac{\pi(2k + 1)}{m}, i, m), k = 0,\dots,m-1 \}
\]
be a set of triples with first element a real number, second element from $D$ and third element from $G$.  Let $H = \abs{T}$ be the number of elements in $T$ and observe that $H \leq \sum_{m\in G}(m+1)\abs{D_m} = O(L)$.  Let $t_1,\dots,t_H$ be an enumeration of the triples from $T$ and denote the three elements of $t_i$ by $a_{i}$, $b_{i}$ and $c_{i}$ respectively, that is,
\[
t_i = ( a_{i}, b_{i}, c_{i} ).
\]
So 
Let $\sigma$ denote the permutation of the integers $\{1,\dots,H\}$ such that $t_{\sigma(i)1}$ are in ascending order, that is,
\begin{equation}\label{eq:sigmasortind}
z_{\sigma_m(i)} \leq z_{\sigma_m(k)}
\end{equation}
whenever $i < k $ where $i, k \in \{0, 1, \dots, \abs{D}-1\}$.  It is convenient to define the indices into $\sigma$ to be taken modulo $\abs{D}$, that is, if $m$ is an integer not from $\{0, 1, \dots, \abs{D}-1\}$ then we define $\sigma(m) = \sigma(k)$ where $k \equiv m \mod \abs{D}$ and $k \in  \{0, 1, \dots, \abs{D}-1\}$.  The first sequence in $S$ is 
\[
f_0 = f(0) = \{ \hat{d}_i(0), i \in D \} = \{ b_i, i \in D \}.
\]  
The next sequence $f_1$ is given by replacing the element $b_{\sigma(1)}$ in $f_0$ with $b_{\sigma(1)}e^{-j2\pi/M}$.  Given a sequence $x$ we use $x e_i$ to denote $x$ with the $i$th element replaced by $x_i e^{-j2\pi/M}$.  Using this notation,  
\[
f_1 = f_0 e_{\sigma(0)}.
\] 
The next sequence in $S$ is correspondingly 
\[
f_2 = f_0 e_{\sigma(0)} e_{\sigma(1)} = f_1 e_{\sigma(1)},
\]
and the $k$th sequence is
\begin{equation}\label{eq:fkrec}
f_{k+1} = f_{k} e_{\sigma(k)}.
\end{equation}
In this way, all $(M+1)\abs{D}$ sequences in $S$ can be recursively enumerated.

We want to find the $f_k \in S$ corresponding to the minimiser of~(\ref{eq:SSdatasymbols}).  A na\"{\i}ve approach would be to compute $SS(f_k)$ for each $k \in \{0,1,\dots,(M+1)\abs{D}-1\}$.  Computing $SS(f_k)$ for any particular $k$ requires $O(L)$ arithmetic operations.  So, this na\"{\i}ve approach would require $O(L (M+1) \abs{D}) = O(L^2)$ operations in total.  Following Mackenthun~\cite{Mackenthun1994}, we show how $SS(f_k)$ can be computed recursively.

Let,
\begin{equation}\label{eq:SSfk}
SS(f_k) = A - \frac{1}{L}\abs{Y_k}^2,
\end{equation}
where, 
\begin{align*}
Y_k = Y( f_k ) &= \sum_{i \in P} y_i p_i^*  + \sum_{i \in D} y_i f_{ki}^* \\
&= B + \sum_{i \in D}g_{ki},
\end{align*}
where $B = \sum_{i \in P} y_i p_i^*$ is a constant, independent of the data symbols, and $f_{ki}$ denotes the $i$th symbol in $f_k$, and for convenience, we put $g_{ki}  = y_i f_{ki}^*$.  Letting $g_{k}$ be the sequence $\{g_{ik}, i \in D\}$ we have, from~\eqref{eq:fkrec}, that $g_k$ satisfies the recursive equation
\[
g_{k+1} = g_{k} e_{\sigma(k)}^*,
\]
where $g_{k} e_{\sigma(k)}^*$ indicates the sequence $g_k$ with the $\sigma(k)$th element replaced by $g_{k \sigma(k)}e^{j2\pi/M}$.  Now,
\[
Y_0 = B + \sum_{i \in D} g_{0i}
\] 
can be computed in $O(L)$ operations, and
\begin{align*}
Y_1 &= B + \sum_{i \in D} g_{1i} \\
&= B +  (e^{j2\pi/M} - 1)g_{0\sigma(0)} + \sum_{i \in D} g_{0i} \\
&= Y_0 + \eta g_{0\sigma(0)},
\end{align*}
where $\eta = e^{j2\pi/M} - 1$.  In general,
\[
Y_{k+1} = Y_k + \eta g_{k\sigma(k)}.
\]
So, each $Y_k$ can be computed from it predecessor $Y_{k-1}$ in a constant number of operations.  Given $Y_k$, the value of $SS(f_k)$ can be computed in a constant number of operations using~\eqref{eq:SSfk}.  Let $\hat{k} = \arg\min SS(f_k)$.  The least squares estimator of the complex amplitude is then computed according to~\eqref{eq:hata},
\begin{equation}\label{eq:ahatYhat}
\hat{a} = \frac{1}{L} Y_{\hat{k}}.
\end{equation}
Psuedocode is given in Algorithm~\ref{alg:loglinear}.  Line~\ref{alg_sortindices} contains the function $\operatorname{sortindicies}$ that, given $z = \{z_i, i \in D\}$, returns the permutation $\sigma$ as described in~\eqref{eq:sigmasortind}.  The $\operatorname{sortindicies}$ function requires sorting $\abs{D}$ elements.  This requires $O(L \log L)$ operations.  The $\operatorname{sortindicies}$ function is the primary bottleneck in this algorithm when $L$ is large.  The loops on lines~\ref{alg_loop_setup} and~\ref{alg_loop_search} and the operations on lines~\ref{alg_Y} to lines~\ref{alg_Q} all require $O(L)$ or less operations.  %In the next sections we will show how to the sortinces function can be avoided.  This leads to an algorithm that requires only $O(L)$ operations.

\begin{algorithm}[t] \label{alg:loglinear}
\SetAlCapFnt{\small}
\SetAlTitleFnt{}
\caption{Mackenthun's algorithm with pilot symbols}
\DontPrintSemicolon
\KwIn{$\{y_i, i \in P \cup D \}$}
\For{$i \in D$ \nllabel{alg_loop_setup}}{
$\phi = \angle{y_i}$ \;
$u = \round{\phi} $ \;
$z_i = \phi -  u $ \;
$g_i = y_i e^{-j u}$ \;
}
$\sigma = \operatorname{sortindicies}(z)$ \nllabel{alg_sortindices} \;
$Y = \sum_{i \in P} y_i p_i^* + \sum_{i \in D} g_i $ \nllabel{alg_Y}\;
$\hat{a} = \frac{1}{L} Y$ \;
$\hat{Q} = \frac{1}{L}\abs{Y}^2$ \nllabel{alg_Q} \;
$\eta = e^{j2\pi/M} - 1$ \;
\For{$k= 0$ \emph{\textbf{to}} $(M+1)\abs{D}-1$ \nllabel{alg_loop_search}}{
$Y = Y + \eta g_{\sigma(k)}$ \;
$g_{\sigma(k)} = (\eta + 1) g_{\sigma(k)} $\;
$Q = \frac{1}{L}\abs{Y}^2$\;
\If{$Q > \hat{Q}$}{
 	$\hat{Q} = Q$ \;
 	$\hat{a} =  \frac{1}{L} Y$ \;
 }
}
\Return{$\hat{a}$ \nllabel{alg_return}}
\end{algorithm}

>>>>

The sequences in $S$ can be enumerated as follows.  Let 
\[
T = \bigcup_{m\in G} \bigcup_{ i \in D_m }\bigcup_{k=1}^{m}\left\{ (z_i + \tfrac{\pi(2k - 1)}{m}, i, m) \right\}
\]
be a set of triples with first element a real number, second element from $D$ and third element from $G$.  Let 
\[
H = \abs{T} = \sum_{m\in G}(m+1)\abs{D_m} = O(L)
\]  
denote the number of elements in $T$.  Let $t_1,\dots,t_H$ be an enumeration of the triples from $T$ sorted in accending order of the first element in the triple, that is, if $t_{i} = (t_{i1},t_{i2},t_{i3})$ then 
\[
t_{i1} \leq t_{ki}
\]
whenever $i < k$.  Put $\sigma(i) = t_{i2}$ and $m(i) = t_{i3}$ so that $\sigma(1), \dots, \sigma(H)$ and $m(1),\dots,m(H)$ correspond respectively with the second and third elements from the triples $t_1,\dots,t_H$ 
%So $a_1,\dots,a_H$ is a sequence of real numbers, $b_1,\dots,b_H$ is a sequence of integers from $D$ and $c_1,\dots,c_H$ is a sequence

The first sequence in $S$ is 
\[
f_1 = f(0) = \{ \hat{d}_i(0), i \in D \} = \{ b_i, i \in D \}.
\]  
The next sequence $f_1$ is given by replacing the element $b_{\sigma(1)}$ in $f_0$ with $b_{\sigma(1)}e^{-j2\pi/m(1)}$.  Given a sequence $x$ we use $x e_{k}$ to denote $x$ with the $\sigma(k)$th element replaced by $x_{\sigma(k)} e^{-j2\pi/m(k)}$.  Using this notation,  
\[
f_2 = f_1 e_1.
\] 
The next sequence in $S$ is correspondingly 
\[
f_3 = f_1 e_{1} e_{2} = f_2 e_{2},
\]
and the $k$th sequence is
\begin{equation}\label{eq:fkrec}
f_{k+1} = f_{k} e_{k}.
\end{equation}
In this way, all $(M+1)\abs{D}$ sequences in $S$ can be recursively enumerated.

We want to find the $f_k \in S$ corresponding to the minimiser of~(\ref{eq:SSdatasymbols}).  A na\"{\i}ve approach would be to compute $SS(f_k)$ for each $k \in \{1,\dots,H\}$.  Computing $SS(f_k)$ for any particular $k$ requires $O(L)$ arithmetic operations.  So, this na\"{\i}ve approach would require $O(LH) = O(L^2)$ operations in total.  Following Mackenthun~\cite{Mackenthun1994}, we show how $SS(f_k)$ can be computed recursively.

Let,
\begin{equation}\label{eq:SSfk}
SS(f_k) = A - \frac{1}{L}\abs{Y_k}^2,
\end{equation}
where, 
\begin{align*}
Y_k = Y( f_k ) &= \sum_{i \in P} y_i p_i^*  + \sum_{i \in D} y_i f_{ki}^* \\
&= B + \sum_{i \in D}g_{ki},
\end{align*}
where $B = \sum_{i \in P} y_i p_i^*$ is a constant, independent of the data symbols, and $f_{ki}$ denotes the $i$th symbol in $f_k$, and for convenience, we put $g_{ki}  = y_i f_{ki}^*$.  Letting $g_{k}$ be the sequence $\{g_{ik}, i \in D\}$ we have, from~\eqref{eq:fkrec}, that $g_k$ satisfies the recursive equation
\[
g_{k+1} = g_{k} e_{k}^*,
\]
where $g_{k} e_{k}^*$ indicates the sequence $g_k$ with the $\sigma(k)$th element replaced by $g_{k \sigma(k)}e^{j2\pi/m(k)}$.  Now,
\[
Y_1 = B + \sum_{i \in D} g_{1i}
\] 
can be computed in $O(L)$ operations, and
\begin{align*}
Y_2 &= B + \sum_{i \in D} g_{1i} \\
&= B +  (e^{j2\pi/m(1)} - 1)g_{1\sigma(1)} + \sum_{i \in D} g_{1i} \\
&= Y_0 + (e^{j2\pi/m(1)} - 1)g_{1\sigma(1)},
\end{align*}
In general,
\[
Y_{k+1} = Y_k + (e^{j2\pi/m(k)} - 1) g_{k\sigma(k)}.
\]
So, each $Y_k$ can be computed from it predecessor $Y_{k-1}$ in a constant number of operations.  Given $Y_k$, the value of $SS(f_k)$ can be computed in a constant number of operations using~\eqref{eq:SSfk}.  Let $\hat{k} = \arg\min SS(f_k)$.  The least squares estimator of the complex amplitude is then computed according to~\eqref{eq:hata},
\begin{equation}\label{eq:ahatYhat}
\hat{a} = \frac{1}{L} Y_{\hat{k}}.
\end{equation}
Psuedocode is given in Algorithm~\ref{alg:loglinear}.  BLERG

%Line~\ref{alg_sortindices} contains the function $\operatorname{sortindicies}$ that, given $z = \{z_i, i \in D\}$, returns the permutation $\sigma$ as described in~\eqref{eq:sigmasortind}.  The $\operatorname{sortindicies}$ function requires sorting $\abs{D}$ elements.  This requires $O(L \log L)$ operations.  The $\operatorname{sortindicies}$ function is the primary bottleneck in this algorithm when $L$ is large.  The loops on lines~\ref{alg_loop_setup} and~\ref{alg_loop_search} and the operations on lines~\ref{alg_Y} to lines~\ref{alg_Q} all require $O(L)$ or less operations.  %In the next sections we will show how to the sortinces function can be avoided.  This leads to an algorithm that requires only $O(L)$ operations.

\begin{algorithm}[t] \label{alg:loglinear}
\SetAlCapFnt{\small}
\SetAlTitleFnt{}
\caption{Mackenthun's algorithm with pilot symbols}
\DontPrintSemicolon
\KwIn{$\{y_i, i \in P \cup D \}$}
\For{$m \in G$ \nllabel{alg_loop_setup}}{
\For{$i \in D_m$ }{
$\phi = \angle{y_i}$ \;
$u = \round{\phi}_m $ \;
$g_i = y_i e^{-j u}$ \;
\For{$k = 1, \dots, m$ }{ $t_{mik} =  (z_i + \tfrac{\pi(2k + 1)}{m}, i, m)$} 
}
}
$\operatorname{sort}(t_1,\dots,t_H)$ \;
$Y = \sum_{i \in P} y_i p_i^* + \sum_{i \in D} g_i $ \nllabel{alg_Y}\;
$\hat{a} = \frac{1}{L} Y$ \;
$\hat{Q} = \frac{1}{L}\abs{Y}^2$ \nllabel{alg_Q} \;
\For{$k = 1,\dots,H$ \nllabel{alg_loop_search}}{
$Y = Y + (e^{j2\pi/m(k)} - 1) g_{\sigma(k)}$ \;
$g_{\sigma(k)} = e^{j2\pi/m(k)} g_{\sigma(k)} $\;
$Q = \frac{1}{L}\abs{Y}^2$\;
\If{$Q > \hat{Q}$}{
 	$\hat{Q} = Q$ \;
 	$\hat{a} =  \frac{1}{L} Y$ \;
 }
}
\Return{$\hat{a}$ \nllabel{alg_return}}
\end{algorithm}
